The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-08-01 carbon_paper/tex/main.tex
tex updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
%referee,
usenatbib]{mnras}

@@

% citations
\defcitealias{james+21}{J21}
\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\textsc{vice}}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.3}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah{} across different environments and metallicities.

@@

We know, from observations of MW stars and (extragalactic) gas, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,%
\footnote{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}
%
observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.
    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep).
    According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced \textit{first dredge up}%
\footnote{First dredge up is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}
yet have well-mixed envelopes. So, subgiant surface abundances most accurately represent their birth composition. 
In contrast, evolved RGB stars require modelling-dependent corrections to recover birth C abundances \citep[e.g.][]{vincenzo+21}. 
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

@@

\section{Nucleosynthesis}
We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

Yields -- the amounts of each chemical element stars synthesise -- are central to studies of galactic chemical evolution (GCE). 
In this section, we compare literature yield predictions and discuss our fiducial yield choices. 
We focus on three nucleosynthetic pathways: asymptotic giant branch (\agb{}) stars, core collapse supernovae (\cc{}), and type Ia supernovae (\ia{}).
C is produced in both \agb{} and \cc{} stars.
We also use Mg and Fe as tracers of \cc{} and \ia{} enrichment respectively. 
O and Mg are produced almost entirely from \cc\ with metallicity-independent yields. In contrast, Fe is produced in similar amounts by \cc\ and \ia.

After a single stellar population (SSP)\footnotemark{} forms, \cc{} are the first chemical enrichers. \cc{} explode within $\lesssim 40$\,Myr, providing light elements (e.g. C, O, and Mg) and heavier elements (Fe and beyond). Next, low-mass stars begin to reach the end of their lives, entering the \agb{} phase. By shedding their outer layers, \agb{} stars are important sources of C, N, and neutron capture elements.  Finally, white dwarfs explode in \ia{}, releasing Fe and other iron-peak elements.

\footnotetext{SSP: Single stellar population. Technical name for a group of stars born in the same conditions at the same time, i.e. an open cluster.}

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} =  \Delta Z_X \frac{M_{\rm ejected}}{M_{\rm birth}}
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in $Z_X$ from the birth material to the ejected material of the star.%
\footnote{$Z_{X}$ represents the mass fraction of element $X$.}
%
For example, a 1\,\Mo\ star with $\y_{\C} = 10^{-3}$ will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Also, note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_X$ than the material the star was formed from.
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, initial-mass-function (\imf) averaged yields are useful in interpreting their predictions. An IMF-averaged yield adds together the yields of stars of each mass, weighted by the fraction of stars of each mass (the \imf). 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE).\footnotemark{}

\addtocounter{footnote}{-2}
    \stepcounter{footnote}\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}.
We use $t_{\rm end}=10\,$Gyr for total yields when $t$ is not used.}

\stepcounter{footnote}\footnotetext{\VICE~is available at \url{https://github.com/giganano/VICE}}

\begin{table}
	\centering
    \caption[]{Yields for the fiducial model (in units of \ssp~birth mass). See Section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $2.3\times$\cxi &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An \agb\ star is a low-mass ($\lesssim 8\,\Mo$) star during its final phase of evolution. Inside an \agb\ star, an inert CO core is surrounded by thermally-unstable He and H-burning shells below a convective envelope. 
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Above \about{2}\,\Mo{}, third dredge up becomes important, enriching the outer layers with C.
\agb\ stars between 2 and 5 \Mo are the most abundant producers of C.
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, hot bottom burning is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.
Variations in models are due to different treatments of reaction rates, convection, and mass-loss.
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{K10}{\texttt{K10}}: \citet{karakas10}
\item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
\end{description}

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity. Metal poor stars dredge up more material due to the decreased power of the CNO cycle, resulting in increasing carbon yields with decreasing metalicity \citep{ventura+13}.

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\y_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass \agb\ stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. Fe production, in contrast, continues steadily for 10\,Gyr.

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, all four models predict $y_\text{C}^\text{agb}(\Zo)$ to be between 0.004 and 0.008 at solar metallicity. The models also differ slightly in the strength of metallicity dependence, with about a factor of 3 discrepency (see Table~\ref{tab:alpha_agb}).

@@

	\label{tab:alpha_agb}
    \begin{tabular}{c ccc p{5cm} p{5cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C}^{\rm agb}(\Zo)$ & $\zeta^{\rm agb}(\Zo)$ & $\alpha_{20}^{\rm agb}$
                & masses (\Mo) & metallicites 
                \\
        \hline
        \cxi 
                &  0.00042 & -0.0175 & 2.4
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \kx 
                & 0.00064 & -0.059 & 1.6
                & 1, 1.25, 1.5, 1.75, 1.9, 2.25, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6 
                &  0.0001, 0.004, 0.008, 0.02
                \\
        \vxiii 
                & 0.00022 & -0.021 & 4.5
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \kxvi 
                & 0.0005 & -0.029 & 2.0
                & 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
		\hline
	\end{tabular}
\end{table*}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process.\footnote{The 3-$\alpha$ process fuses 3 He nuclei into \C[12].} 
However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

Fig.~\ref{fig:y_cc} plots calculations of the \imf-integrated yields, defined with Eq.~\ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). 
\cc{} models predict a wide range of C yields, spanning almost a factor of ten. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
As metallicity increases, stars lose more of their mass to winds. In particular, C enriched envelop material is lost through winds before synthesized into heavier elements, so C yields can be strongly metallicity dependent (VERIFY).
Fig.~\ref{fig:y_cc} shows the \cxi{} \agb\ model for comparison on the left. Especially at $Z\approx Z_\odot$, most \cc{} models dominate \agb\ C production. Later, we will also show empirically this is the case.
The right of Fig.~\ref{fig:y_cc} shows the \cc{} [C/Mg] ratio for the different models, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
If only \cc{} produced C, then ${\rm [C/Mg]^{CC}}$ describes the equilibrium abundance of [C/Mg].
Different \cc\ models also span a large range in [C/Mg]. 
We chose to instead parameterize $\Ycc$ to simplify model inputs, as most \cc\ models fail to achieve near-solar [C/Mg].

Both rotation and explodability introduce substantial variations. The \cite{LC18} models include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger and contributes to wind losses. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.
Assumptions about the explodability landscape affect C and Mg production. As fewer high-mass stars explode, both C and Mg yields decrease, but Mg yields decrease more as more C is lost to winds, so [C/Mg] increases with decreasing explodability.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
% referee,
usenatbib]{mnras}

@@

% citations
\defcitealias{james+21}{J21}
\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\textsc{vice}}

\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.4}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a single-zone model with our recommended yields replicate the broad trends of \caah{} across different environments and metallicities.

@@

We know, from observations of MW stars and (extragalactic) gas, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,%
\footnote{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}
%
observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.
    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep).
According to stellar evolution theory and observations, subgiants most accurately represent their birth composition \citep{gilroy89, korn+07, lind+08, souto+18, souto19}.
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

@@

Subgiants provide the ideal observational constraint to our model. 
When a star enters the Red Giant Branch (RGB), material from the CNO-processed core is mixed with the envelope in first dredge up, enhancing N and depleting C \citep{iben67, vincenzo+21,KL14}. RGB stars thus require model-dependent corrections to recover surface abundances \citep[e.g.][]{vincenzo+21}. On the other hand, gravitational settling can affect main sequence stellar abundances \citep[e.g.][]{souto19}. Subgiants have well-mixed envelopes, so gravitational settling is not as significant, and subgiants have not yet experienced first dredge up. We use a sample of \apogee\ DR17 stars \citep{apogee17} as selected in \citetjack.  
Chemical abundances are determined from the \apogee\ Stellar Parameter and Chemical Abundance Pipeline ({\sc aspcap}) \citep{aspcap}.

Fig.~\ref{fig:subgiant_selection} shows a plot of all \apogee\ stars and the \citetjack polygon selection criteria. 
 \citetjack~select a region of stars based on surface gravity $\log g$, and effective surface temperature, $T_\text{eff}$.
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.000706\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Additionally, we exclude stars in \apogee{} marked by any of the following flags.
\begin{description}
\item \verb|APOGEE_MIRCLUSTER_STAR|
\item \verb|APOGEE_EMISSION_STAR|
\item \verb|APOGEE_EMBEDDEDCLUSTER_STAR|
\item \verb|young cluster (IN-SYNC)|
\item \verb|APOGEE2_W345|
\item \verb|EB planet|
\end{description}
Our final sample contains \about{12,000} subgiants.

When plotting the data and models, we bin by groups of 500 (or 150) stars in [Mg/H] ([Mg/Fe]). When plotting \caah, we select only the low-$\alpha$ sequence. For \caafe, we instead select stars with $-0.15 \leq {\rm [Mg/H]} \leq -0.05$.  
When plotting \caah, we isolate the low- and high-$\alpha$ sequences with the cut
\begin{equation}\label{eq:high_alpha}
\begin{cases}
\text{[Mg/Fe]} >0.12-0.13\,\text{[Fe/H]}, & \text{[Fe/H]}<0\\
\text{[Mg/Fe]} >0.12, & \text{[Fe/H]}>0. \\
\end{cases}
\end{equation}
The low-$\alpha$ sequence is better reproduced by this model, so we use this cut of the subgiants to compare the models against except for comparing \caafe.

\section{Nucleosynthesis}
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
For O, N, Mg, and Fe, we adapt the yield choices of \citet{james+21, james+23}.
Following \citet{james+21, james+23}, we take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

\begin{table}
	\centering
    \caption[]{Yields for the fiducial model (in units of \ssp~birth mass). See section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $\cfactor\times$\cxi &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{K10}{\texttt{K10}}: \citet{karakas10}
\item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
\end{description}
Table~\ref{tab:agb} contains the masses and metallicities used in each model.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
We define the stellar \agb\ yield to be the fraction of the stars initial mass $M$ which is newly synthesized into C. If $\Delta Z_{\rm C}$ represents the change in the mass fraction of C abundance, then
\begin{equation}
    \y_{\rm C}^{\rm agb} =  \Delta Z_{\rm C} \frac{M_{\rm ejected}}{M}.
\end{equation}
Note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_{\rm C}$ than the material the star was formed from.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity.

The left panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
An IMF-averaged yield adds together the yields of stars of each mass, weighted by the fraction of stars of each mass (the \imf). 
\begin{equation} \label{eq:imf-agb-yield}
    \Ycagb(Z,t) = 
    \frac{
    \int_{M_{\rm to}(t)}^{8\,\Mo} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int_{0.08\Mo}^{100\Mo} \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf{} and $M_{\rm to}(t)$ is the mass of stars with lifetime $t$.%
\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}.
We use $t_{\rm end}=10\,$Gyr for total yields when $t$ is not used.}
The \agb\ models mostly differ in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, all four models predict $y_\text{C}^\text{agb}(\Zo)$ to be between 0.004 and 0.008 at solar metallicity. The models also differ slightly in the strength of metallicity dependence, with about a factor of 3 discrepency (see Table~\ref{tab:agb}).

Fig.~\ref{fig:agb-ssp}, on the right, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\y_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass \agb\ stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. Fe production, in contrast, continues steadily for 10\,Gyr.

In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass and metallicity dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Above \about{2}\,\Mo{}, third dredge up becomes important, enriching the outer layers with C.
\agb\ stars between 2 and 5 \Mo are the most abundant producers of C.
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, hot bottom burning is much more efficient, resulting in significant \C[12] destruction.
Metal poor stars dredge up more material due to the decreased power of the CNO cycle, resulting in increasing carbon yields with decreasing metalicity \citep{ventura+13}.

For our models to match observations, we find that need to uniformly amplify these yield tables.
\begin{equation} \label{eq:alpha}
        \Ycagb \rightarrow \alpha_{\rm agb}\ \Ycagb.
\end{equation}
We use \cxi\ table, with $\alpha_{\rm agb}=\cfactor$, as the fiducial \agb\ yield.
Variations in models are due to different treatments of reaction rates, convection, and mass-loss.
Table~\ref{tab:agb} contains the stellar masses and metallicites each yield set contains, the approximate solar C yield $\Ycagb(\Zo)$, the metallicity dependence at solar $\zeta^{\rm agb} = d\Ycagb/dZ$, and the required factor $\alpha$ to reach 20\% \agb\ C production.

@@

	\label{tab:agb}
    \begin{tabular}{c ccc p{5cm} p{5cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C}^{\rm agb}(\Zo)$ & $\zeta^{\rm agb}(\Zo)$ & $\alpha_{20}^{\rm agb}$
                & masses (\Mo) & metallicites 
                \\
        \hline
        \cxi 
                &  0.00042 & -0.0175 & 2.4
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \kx 
                & 0.00064 & -0.059 & 1.6
                & 1, 1.25, 1.5, 1.75, 1.9, 2.25, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6 
                &  0.0001, 0.004, 0.008, 0.02
                \\
        \vxiii 
                & 0.00022 & -0.021 & 4.5
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \kxvi 
                & 0.0005 & -0.029 & 2.0
                & 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
		\hline
	\end{tabular}
\end{table*}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process.
However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

Fig.~\ref{fig:y_cc} plots calculations of the \imf-integrated yields.
For \cc\, 
\begin{equation} \label{eq:imf-cc-yield}
    \Ycc(Z) = 
    \frac{
    \int_{8\,\Mo}^{100\,\Mo} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int_{0.08\Mo}^{100\Mo} \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf{} and $M_{\rm to}(t)$ is the mass of stars with lifetime $t$.%
\cc{} models predict a wide range of C yields, spanning almost a factor of ten. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
As metallicity increases, stars lose more of their mass to winds. In particular, C enriched envelop material is lost through winds before synthesized into heavier elements, so C yields can be strongly metallicity dependent (VERIFY).
Fig.~\ref{fig:y_cc} shows the \cxi{} \agb\ model for comparison on the left. Especially at $Z\approx Z_\odot$, most \cc{} models dominate \agb\ C production. Later, we will also show empirically this is the case.
The right of Fig.~\ref{fig:y_cc} shows the \cc{} [C/Mg] ratio for the different models, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
If only \cc{} produced C, then ${\rm [C/Mg]^{CC}}$ describes the equilibrium abundance of [C/Mg].
Different \cc\ models also span a large range in [C/Mg].

Rotation, binarity, and explodability introduce substantial uncertainties in \cc\ predictions. The \cite{LC18} models include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger and contributes to wind losses. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.
Assumptions about the explodability landscape affect C and Mg production. As fewer high-mass stars explode, both C and Mg yields decrease, but Mg yields decrease more as more C is lost to winds, so [C/Mg] increases with decreasing explodability. 
Finally, stripped binaries produce about twice as much C as single massive stars and other effects of binary evolution are unstudied \citep{farmer+21}.

\cc\ do not satisfactorily explain C, O, and Mg abundances.
Most models here overproduce C relative to Mg. This could be related to overproduction of O relative to Mg, a known problem of \cc\ models.
There is substantial variation in predicted Mg yields but most models predict flat trends. However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models, but O and Mg yields of \cc\ models do not fully match observations. \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown (see e.g. \citealt{emily+21}). Here, we assume ${\rm [O/Mg]} = 0$, which nevertheless consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.
 As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text, leaving them constant and metallicity-independent. We also choose to parameterize C yields.
